---
title: "linear models"
author: "ASHLEY ROMO"
date: "2023-11-18"
output: github_document
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Load key packages.

```{r}
library(tidyverse)
library(p8105.datasets)
```



## Load and clean the Airbnb data

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  select(
    price, stars, borough = neighbourhood_group, 
    neighbourhood, room_type) |> 
  filter(borough != "Staten Island")

```


Let's fit a model!!

```{r}
# lm(price ~ stars + borough, date = nyc_airbnb)
fit = 
  nyc_airbnb |> 
  lm(price ~ stars + borough, data = _) #underscore here specifies the dataset
```


Let's look at the `fit`

```{r}
fit # not very helpful

summary(fit) #provides residuals info and coefficient info (estimate, sd error, t value, and p value)

summary(fit)$coef #take out the coefficient table 

coef(fit) # take out value of coefficients alone

# fitted.values(fit) #take out all of the fitted values
```

tidy up the output instead

```{r}
#better way to do this is using glance()

#glance() extracts general model fits statistics aka high level summary of how well the model works, including R^2, adjusted R^2, overal gloval p value, etc

fit |> 
  broom::glance()
```


tidy up the coefficients

```{r}
#use the tidy() function pulls out the coefficient table

#tidy() is a tibble (better than glance because glance outputs a matrix)

fit |> 
  broom::tidy() |> 
  mutate(term = str_replace(term, "^borough", "Borough: ")) |> 
  select(term, estimate, p.value) |> 
  knitr::kable(digits = 3)
```

## Fit another model


```{r}
fit = 
  nyc_airbnb |>
lm(price ~ stars + borough + room_type, data = _)
# when R is creating these indicator variables, it is taking the variable you started out with and assuming there is a factor ordering in alphabetical order
#here bronx is the reference category because it is the first alphabetically, but you can specificy the order


fit = 
  nyc_airbnb |> 
  mutate(
    borough = fct_infreq(borough), #factor based on frequency, so reference category changes. Now, reference is Manhattan
    room_type = fct_infreq(room_type)
  ) |> 
  lm(price ~ stars + borough + room_type, data = _)

fit |> 
  broom::tidy()
```



## Quick look at diagnostics (minute 32)

```{r}
#regression diagnostics boils down to get the residuals and do a careful inspection of residuals to make sure you don't have overly skewed distributions and non-normal distributions
#make sure you have the residuals you meed by using add_residuals(dataframe,added residuals form the model you just fit) function
#output gives us a residual column that shows what happens after you fit this model (here, it is the model = fit) and apply it to this dataset (Data = nyc_airbnb) to compute the residuals that go along with the data

#you can also use this function to check if your model is generalizable  
nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(x = resid)) + #helpful first look
  geom_density() +
  xlim(-100, 500)


nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(x = borough, y = resid)) +  #can plot your residuals
  geom_violin() 

nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(x = stars, y = resid)) + 
  geom_point() 

##always need to check residuals to make sure we are not violating the assumption 
```

Now with a qq plot!

```{r}
nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(sample = resid)) + 
  stat_qq() + 
  stat_qq_line()
```

## Hypothesis test for categorical predictor

fit a "null" and "alternative" model

```{r}
fit |> 
  broom::tidy() #this gives the p values for each variable to see if it is significant

#if you wanted to ask if you need room type (yes vs no), you need to compare a model with the categorical predictor and one without 
fit_null = lm(price ~ stars + borough, data = nyc_airbnb) #linear model of price on stars and borough using nyc_airbnb model
fit_alternative = lm(price ~ stars + borough + room_type, data = nyc_airbnb)
#alternative is the null plus room type  (we want to know if room type is significant)

#can also use the ANOVA 
anova(fit_null, fit_alternative) |> 
  broom::tidy() 
```

## Borough-level differences

```{r}
fit = 
  nyc_airbnb |> 
  lm(price ~ stars * borough + room_type * borough, data = _) #asking whether or not price depends on stars and room type in this data set but want to allow  the effect of stars to vary by boroughs and the effect of room type vary by borough

fit |> 
  broom::tidy() # this gives you all the estimates and p-values
```



```{r}
#for exploratory analyses, you can fit a separate linear model for each borough and look at the slope of stars

airbnb_lm = function(df) {
  lm(price ~ stars + room_type, data = df)
}

#nest everything except for borough mutate( #fit a linear model for each borough; models column is going to map across the dataframe column and im going to fit a linear model that goes along with that
nyc_airbnb |> 
  nest(df = -borough) |> 
  mutate(
    models = map(df, airbnb_lm),
    results = map(models, broom::tidy)    
  ) |> 
  select(borough, results) |> 
  unnest(results) |> 
  select(borough, term,  estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |> 
  knitr::kable(digits = 2)
```

same thing but just a little different

```{r}
# \() anonymous function

nyc_airbnb |> 
  nest(df = -borough) |>  
    mutate(
      models = map(df, \(df) lm(price ~ stars + room_type, data = df)),
      results = map(models, broom::tidy)
  ) |> 
  select(borough, results) |> 
  unnest(results) |> 
  select(borough, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |> 
  knitr::kable(digits = 2)
```

## Homicides in Baltimore

```{r}
# disposition == "Closed by arrest") results in true or false. Using as.numeric in front of it converts it to 1 if true and 0 if false 
baltimore_df = 
  read_csv("data/homicide-data.csv") |> 
  filter(city == "Baltimore") |> 
  mutate(
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age)
  ) |> 
  select(resolved, victim_age, victim_race, victim_sex)
```


fitting a logistic regression

```{r}
fit_logistic = 
  baltimore_df |> 
  glm(
    resolved ~ victim_age + victim_race + victim_sex, 
    data = _, 
    family = binomial())
```

look at model results

```{r}
fit_logistic |> 
  broom::tidy() |> 
  mutate(OR = exp(estimate)) |> 
  select(term, estimate, OR)
```


```{r}
baltimore_df |> 
  count(victim_race)
```

